Best Practices for Single Alchemical Free Energy Calculations [Article v\versionnumber]

### 2025-06-10 paper/manuscript.tex

\newcommand{\versionnumber}{2.0} % you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/alchemistry/alchemical-best-practices}} %this should be the main github repository for this article
\newcommand{\expect}[1]{\left\langle{#1}\right\rangle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Single Alchemical Free Energy Calculations [Article v\versionnumber]}
\author[1*]{Antonia S. J. S. Mey}
\author[2]{Bryce K. Allen}
\author[3]{Hannah E. Bruce Macdonald}
\author[3*]{John D. Chodera}
\author[9]{David F. Hahn}
\author[1,10]{Maximilian Kuhn}
\author[1]{Julien Michel}
\author[4*]{David L. Mobley}
\author[5]{Levi N. Naden}
\author[6]{Samarjeet Prasad}
\author[2,7]{Andrea Rizzi}
\author[1]{Jenke Scheen}
\author[8*]{Michael R. Shirts}
\author[9]{Gary Tresadern}
\author[2]{Huafeng Xu}
%
\affil[1]{EaStCHEM School of Chemistry, David Brewster Road, Joseph Black Building, The King's Buildings, Edinburgh, EH9 3FJ, UK}
\affil[2]{Silicon Therapeutics, Boston, MA, USA}
\affil[3]{Computational and Systems Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York NY, USA}
\affil[4]{Departments of Pharmaceutical Sciences and Chemistry, University of California, Irvine, Irvine, USA}
\affil[5]{Molecular Sciences Software Institute, Blacksburg VA, USA}
\affil[6]{National Institutes of Health, Bethesda, MD, USA}
\affil[7]{Tri-Institutional Training Program in Computational Biology and Medicine, New York, NY, USA}
\affil[8]{University of Colorado Boulder, Boulder, CO, USA}
\affil[9]{Computational Chemistry, Janssen Research \& Development, Turnhoutseweg 30, Beerse B-2340, Belgium}
\affil[10]{Cresset, Cambridgeshire, UK}

@@

\orcid{Antonia S. J. S. Mey}{0000-0001-7512-5252}
\orcid{Bryce Allen}{0000-0002-0804-8127}
\orcid{Hannah E. Bruce Macdonald}{0000-0002-5562-6866}
\orcid{John D. Chodera}{0000-0003-0542-119X}
\orcid{Maximilian Kuhn}{0000-0002-2811-3934}
\orcid{Julien Michel}{0000-0003-0360-1760}
\orcid{David L. Mobley}{0000-0002-1083-5533}
\orcid{Levi N. Naden}{0000-0002-3692-5027}
\orcid{Samarjeet Prasad}{0000-0001-8320-6482}
\orcid{Andrea Rizzi}{0000-0001-7693-2013}
\orcid{Jenke Scheen}{0000-0001-9781-0445}
\orcid{Michael R. Shirts}{0000-0003-3249-1097}
\orcid{Gary Tresadern}{0000-0002-4801-1644}
\orcid{Huafeng Xu}{0000-0001-5447-0452}
\orcid{David F. Hahn}{0000-0003-2830-6880}

@@

>>>>

\newcommand{\versionnumber}{2.0} % you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/alchemistry/alchemical-best-practices}} %this should be the main github repository for this article
\newcommand{\expect}[1]{\left\langle{#1}\right\rangle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Single Alchemical Free Energy Calculations [Article v\versionnumber]}
\author[1*]{Antonia S. J. S. Mey}
\author[2]{Bryce K. Allen}
\author[3]{Hannah E. Bruce Macdonald}
\author[3*]{John D. Chodera}
\author[9]{Vytautas Gapsys}
\author[9]{David F. Hahn}
\author[1,10]{Maximilian Kuhn}
\author[1]{Julien Michel}
\author[4*]{David L. Mobley}
\author[5]{Levi N. Naden}
\author[6]{Samarjeet Prasad}
\author[2,7]{Andrea Rizzi}
\author[1]{Jenke Scheen}
\author[8*]{Michael R. Shirts}
\author[9]{Gary Tresadern}
\author[2]{Huafeng Xu}
%
\affil[1]{EaStCHEM School of Chemistry, David Brewster Road, Joseph Black Building, The King's Buildings, Edinburgh, EH9 3FJ, UK}
\affil[2]{Silicon Therapeutics, Boston, MA, USA}
\affil[3]{Computational and Systems Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York NY, USA}
\affil[4]{Departments of Pharmaceutical Sciences and Chemistry, University of California, Irvine, Irvine, USA}
\affil[5]{Molecular Sciences Software Institute, Blacksburg VA, USA}
\affil[6]{National Institutes of Health, Bethesda, MD, USA}
\affil[7]{Tri-Institutional Training Program in Computational Biology and Medicine, New York, NY, USA}
\affil[8]{University of Colorado Boulder, Boulder, CO, USA}
\affil[9]{Computational Chemistry, Janssen Research \& Development, Turnhoutseweg 30, Beerse B-2340, Belgium}
\affil[10]{Cresset, Cambridgeshire, UK}

@@

\orcid{Antonia S. J. S. Mey}{0000-0001-7512-5252}
\orcid{Bryce Allen}{0000-0002-0804-8127}
\orcid{Hannah E. Bruce Macdonald}{0000-0002-5562-6866}
\orcid{John D. Chodera}{0000-0003-0542-119X}
\orcid{Maximilian Kuhn}{0000-0002-2811-3934}
\orcid{Julien Michel}{0000-0003-0360-1760}
\orcid{David L. Mobley}{0000-0002-1083-5533}
\orcid{Levi N. Naden}{0000-0002-3692-5027}
\orcid{Samarjeet Prasad}{0000-0001-8320-6482}
\orcid{Andrea Rizzi}{0000-0001-7693-2013}
\orcid{Jenke Scheen}{0000-0001-9781-0445}
\orcid{Michael R. Shirts}{0000-0003-3249-1097}
\orcid{Gary Tresadern}{0000-0002-4801-1644}
\orcid{Huafeng Xu}{0000-0001-5447-0452}
\orcid{David F. Hahn}{0000-0003-2830-6880}
\orcid{Vytautas Gapsys}{0000-0002-6761-7780}

@@

Alchemical Free Energy (AFE) calculations have primarily been applied to simulate the stoichiometric, non-reversible binding of organic molecules to well-structured proteins. However, several areas of drug discovery involve other classes of binding events. While the underlying theory of alchemical free energy calculations is generally applicable, additional work may be required to develop efficient protocols suited to systems beyond the traditional small molecule–protein complex.
For example, there is growing interest in identifying small molecules that directly bind to RNA or DNA, with recent preliminary studies showing promising results using specialized simulation protocols.\cite{Abramyan2025,Rasouli2024} In contrast, there has been limited investigation into the applicability of AFE methods to small-molecule ligands targeting intrinsically disordered proteins (IDPs,\cite{Papadourakis2024} or proteins containing intrinsically disordered regions (IDRs) that undergo disorder-to-order transitions upon ligand binding.\cite{MendozaMartinez2022}
The treatment of covalent binders has also received some attention.\cite{Luo2021,Zhang2019 } However, robust protocols for handling datasets of irreversible covalent binders containing diverse ‘warheads’ have not yet achieved widespread adoption.\cite{Yu2019}
